Identify max rank for AUCell

### 2025-01-16 analyses/cell-type-ewings/exploratory_analysis/07-identify-max-rank-aucell.Rmd
add missing link
Update analyses/cell-type-ewings/exploratory_analysis/07-identify-max-rank-aucell.Rmd
Co-authored-by: Stephanie Spielman <stephanie.spielman@gmail.com>
make the report pretty

From the  [`AUCell` vignette]():

@@

Here we will build the rankings and look at the histogram showing the number of genes detected per cell for each sample and use that information to pick a threshold (percentage of genes) to use across all samples.

@@

    genes_to_remove <- rowData(sce)$detected > 0 
    filtered_sce <- sce[rowData(sce)$detected > 0 , ]

@@

## `AUCell` build rankings

Below we will look at the distribution of the number of genes detected per cell for each sample using the `AUCell_buildRankings()` function.

```{r}
cell_rankings_list <- sce_list |> 
  purrr::map(\(sce){AUCell::AUCell_buildRankings(counts(sce))})
```

@@

```{r}
sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.05)})
```

@@

```{r}
sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.01)})
```

>>>>

From the  [`AUCell` vignette](https://bioconductor.org/packages/release/bioc/vignettes/AUCell/inst/doc/AUCell.html):

@@

Here we will look at a histogram showing the number of genes detected per cell for each sample and use that information to pick a threshold (percentage of genes) to use across all samples.

@@

    genes_to_remove <- rowData(sce)$detected > 0 
    filtered_sce <- sce[genes_to_remove > 0 , ]

@@

## Number of genes detected

Below we will look at the distribution of the number of genes detected per cell for each sample.

```{r, fig.height = 10, fig.width=10, warning=FALSE}
# make a list of the `countByCell` as AUCell does
names(sce_list) <- stringr::str_extract(sce_files, "SCPCS\\d+")
count_df_list <- sce_list |>
  purrr::map(counts) |>
  purrr::map(
    \(counts) {
      countByCell <- Matrix::colSums(counts > 0, na.rm = TRUE)
      data.frame(count_by_cell = countByCell)
    })

# first, make the plots
count_df_list |>
  purrr::imap(
    \(count_df, sample_id) {
        count_hist <- ggplot(count_df) + 
          aes(x = count_by_cell) + 
          geom_histogram(color = "grey20", fill = "skyblue", bins = 30) + 
          xlim(0, max(count_df$count_by_cell)) +
          xlab("Number of genes detected per cell")

        count_box <- ggplot(count_df) + 
          aes(x = count_by_cell, y = "") + 
          geom_boxplot(color = "grey20", fill = "skyblue", outlier.size = 0.5) + 
          theme_void() + 
          xlim(0, max(count_df$count_by_cell)) + 
          ggtitle(sample_id)

      count_box / count_hist + patchwork::plot_layout(heights = c(0.2, 1))
    }
  ) |>
  patchwork::wrap_plots(nrow = 5)

# second, print a table of the stats
count_df_list |>
  purrr::map(
     \(count_df) {
         c(
          min=min(count_df$count_by_cell), 
          quantile(count_df$count_by_cell, c(.01,.05, .10, .50, 1))
        )
     }) |>
  dplyr::bind_rows(.id = "sample_id")
```

@@

```{r}
# get max rank using 1% threshold
threshold_list <- sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.05)})

# make it legible
data.frame(
  max_rank = threshold_list
)
```

@@

```{r}
# get max rank using 1% threshold
threshold_list <- sce_list |> 
  purrr::map_dbl(\(sce){ceiling(nrow(counts(sce))*0.01)})

# make it legible
data.frame(
  max_rank = threshold_list
)
```
